---
title: "Prep Activity 9-21-2021"
author: "Brandon Ritchie"
date: "9/21/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1.

Let $f(x) = \sqrt{x}$ 

$f(1)$
$f(0)$
$f(36)$
$f(1/4)$
$f(15)$
$f(-1)$
```{r, message = FALSE}
fun1 <- function(x){
  x^0.5
}

fun1(1)
fun1(0)
fun1(36)
fun1(0.25)
fun1(15)
fun1(-1)
```

2.

$f(x) = 16$
$f(x) = 1/25$
$f(x) = -1$
```{r}
fun2 <- function(x){
  ifelse(x>0,x^2,NA)
}
fun2(16)
fun2(1/25)
fun2(-1)
```

3.Evaluate $f(x+7)$. Identify the transformation. What does this tell you about the relationship between the graphs of $f(x)$ and $f(x+7)$?

$f(x+7)$ adds 7 to the x variable being passed into the function.

4. Evaluate $f(x) + 3$. Identify the transformation. What does this tell you about the relationship between the graphs of $f(x)$ and $f(x) + 3$?

$f(x) + 3$ translates the function 3 units on the y-axis

5. Given a function $h(t)$, what is the difference between the symbols $h(t) = 5$ and $h(5)$?

$h(t) = 5$ Means that for any value of t passed into the function, the output will always be 5. $h(5)$ means that the value 5 is being passed into the function h.
